Continous time Markov chain models; \\ extension, simulation and visualisation

### 2015-02-11 Response.tex

\begin{quote}
\myeditsveighteen
\end{quote}

% 19
\item {
{\it
\textbf{
P85. Riemannian -> Riemann
}% END: textbf
}% END: it
}% END: item
\\
Corrected.

% 20
\item {
{\it
\textbf{
P86. The last sentence of Section 5.1.3 is not sufficiently clear. 
Please make clear why the use of mixed effects models could allow this.
}% END: textbf
}% END: it
}% END: item
\\
I changed the relevant subsection, explaining that by this sentence we mean the impact of various covariates as well as that by using MG94 parametrization we get a better resolution on selection parameters.
There is also a more elaborate description of the model we plan to use in the upcoming manuscript on the application of the hierarchical epoch model.

\begin{quote}
\myeditsvtwenty
\end{quote}

% 21
\item {
{\it
\textbf{
P88. It would be useful to receive a bit more detail what is meant by `infinite hidden Markov Model'.
}% END: textbf
}% END: it
}% END: item
\\
I introduce the definition of HMM's in a form of Remark on page 93,
% ~\pageref{ref:infinite_hidden_markov}, 
please see a relevant subsection.

@@

\begin{itemize}

% 1
\item {
{\it
\textbf{
I missed a brief primer of Bayesian model selection in the introductory chapter.
Techniques such Bayes factors, path sampling or stepping-stone sampling are used throughout but never explained. 
This should be very easy to fix.
}% END: textbf
}% END: it
}% END: item
\\
There is now a subsection called \emph{Bayesian model selection} at the end of Chapter 1, which gives a primer on Bayesian model selection that should be sufficient to follow the rest of the manuscript.

% 2
\item {
{\it
\textbf{
In the simulations in Chapter 2 you assume you know the true epoch times, but this information is rarely available in real life. 
Simulation 1 addresses the impact of model misspecification partially, as the true process does not have epochs. 
It could be a good idea to further check the impact of model misspecification, in particular simulating a number of epochs that do not correspond with the assumed ones (you later mention this idea in the final chapter but do not end up testing its potential biasing effect). 
Also, simulating under a non-epoch model - for example where different process can apply to contemporaneous branches, would be fundamental to understand the robustness of the model. 
This does not need to be done before the defense, but it could be a nice extension for the future.
}% END: textbf
}% END: it
}% END: item
\\
We are performing a set of simulations to asses this, and the obrained results will be presented during the public defence.

% 3
\item {
{\it
\textbf{
In general, there is some repetitive text, sometimes with identical phrases/paragraphs, between the introduction and some of the other chapters. 
On the other hand, some relevant aspects of the research are discussed in the general discussion in Chapter 5, instead of in the corresponding `research' chapter. 
This is true in particular for Chapter 3, were a very concise description of SPREAD is provided. 
I understand these redundancies and fragmentations are a byproduct of reproducing papers as chapters, but however these issues have been corrected in some occasions. 
For example, chapter 2 is not identical to the SysBio paper. 
I suggest you try a bit to further consolidate the different chapters.
}% END: textbf
}% END: it
}% END: item
\\
We carefully screened for repeated paragraphs between the introduction and the other chapters, and could only identify one paragraph that is repeated at the beginning of Subsection~\ref{sub:subst_models}.
This paragraph is now re-sentenced to remove the repetition.

We acknowledge that isolated sentence might still carry some degree of resemblance, yet we preffer not to remove this in order to also allow the subsequent chapter to function as self-contained work, for readers possibly interested only in a part of the presented work.

### 2015-02-11 chapters/chapter06.tex
fixing floats
implementing comments
citations in resonse

Specifically we parametrize each epoch specific, possibly multidimensional parameter $\Theta_{i},\; i=1,\ldots,M$ as a log-linear function of $P$ predictors $\mathbf{X}=\left(\mathbf{x_1},\ldots,\mathbf{x_P} \right)$:

@@

\begin{remark}{Infinite hidden Markov models}
Hidden Markov model (HMM) $\left\{ X\left(t\right):\; t=1,2,\ldots\right\}$ defines a probability distribution over sequences of observations $\mathbf{X}=\left\{ X(1), \ldots,X(t) \right\}$ by invoking another sequence of unobserved (\emph{hidden}) discrete state variables $\left\{ C\left(t\right):\; t=1,2,\ldots\right\}$.

@@

\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{lemuria}
\caption{
{ \footnotesize 
{\bf Hypothetical Sketch of the Monophyletic Origin and the Extension of the 12 Races of Man from Lemuria over the Earth. E. Haeckel, 1876.} 
} % END: footnotesize
}
\label{fig:lemuria}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[scale=0.3]{tanglemaps}
\caption{
{ \footnotesize 
{\bf Tangle-map of the biodiversity of 19 marine metagenomes.} 
The figure is plotted using GenGis (\url{http://kiwi.cs.dal.ca/GenGIS/Main_Page}).
} % END: footnotesize
}
\label{fig:tanglemaps}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[scale=0.2]{continuous.pdf}
\caption{
{ \footnotesize 
{\bf Visualizing continuous phylogeographic reconstructions with uncertainty.} 
A. Example visualization in Google Earth based in estimates obtained from a real data set. B. Conceptual representation of visualizing uncertainty by time slicing a posterior distribution.
} % END: footnotesize
}
\label{fig:cont}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[scale=0.4]{discrete}
\caption{
{ \footnotesize 
{\bf Phylogeographic visualization of viral spread in discrete space.} 
A. Conceptual representation of time-slicing a phylogeny and projecting the part of the tree up to time $t_{1}$ and $t_{2}$ respectively. 
B. An example projection of influenza H5N1 over time in Google Earth for two different segment phylogenies. % (hemagglutinin and neuraminidase in green and magenta respectively).
} % END: footnotesize
}
\label{fig:discrete}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[scale=0.2]{moore}
\caption{
{ \footnotesize 
{\bf Performance gains of a numerical code benchmarked on a single CPU core over the course of 15 years (PDE solver FeatFlow, \citep{Turek1999}).}
Last two years have seen only a marginal speed-up, and correspond to the shift to multicore architectures.
} % END: footnotesize
}
\label{fig:moore}
\end{figure}

>>>>

\myedit{svtwenty}{
Specifically we parametrize each epoch specific, possibly multidimensional parameter $\Theta_{i},\; i=1,\ldots,M$ as a log-linear function of $P$ predictors $\mathbf{X}=\left(\mathbf{x_1},\ldots,\mathbf{x_P} \right)$:
}

@@

\begin{remark}{Infinite hidden Markov models\label{ref:infinite_hidden_markov}}
Hidden Markov model (HMM) $\left\{ X\left(t\right):\; t=1,2,\ldots\right\}$ defines a probability distribution over sequences of observations $\mathbf{X}=\left\{ X(1), \ldots,X(t) \right\}$ by invoking another sequence of unobserved (\emph{hidden}) discrete state variables $\left\{ C\left(t\right):\; t=1,2,\ldots\right\}$.

@@

\begin{figure}[h!]
\centering
\includegraphics[scale=0.25]{lemuria}
\caption{
{ \footnotesize 
{\bf Hypothetical Sketch of the Monophyletic Origin and the Extension of the 12 Races of Man from Lemuria over the Earth. E. Haeckel, 1876.} 
} % END: footnotesize
}
\label{fig:lemuria}
\end{figure}

@@

\begin{figure}[h!]
\centering
\includegraphics[scale=0.3]{tanglemaps}
\caption{
{ \footnotesize 
{\bf Tangle-map of the biodiversity of 19 marine metagenomes.} 
The figure is plotted using GenGis (\url{http://kiwi.cs.dal.ca/GenGIS/Main_Page}).
} % END: footnotesize
}
\label{fig:tanglemaps}
\end{figure}

@@

\begin{figure}[h!]
\centering
\includegraphics[scale=0.2]{continuous.pdf}
\caption{
{ \footnotesize 
{\bf Visualizing continuous phylogeographic reconstructions with uncertainty.} 
A. Example visualization in Google Earth based in estimates obtained from a real data set. B. Conceptual representation of visualizing uncertainty by time slicing a posterior distribution.
} % END: footnotesize
}
\label{fig:cont}
\end{figure}

@@

\begin{figure}[h!]
\centering
\includegraphics[scale=0.4]{discrete}
\caption{
{ \footnotesize 
{\bf Phylogeographic visualization of viral spread in discrete space.} 
A. Conceptual representation of time-slicing a phylogeny and projecting the part of the tree up to time $t_{1}$ and $t_{2}$ respectively. 
B. An example projection of influenza H5N1 over time in Google Earth for two different segment phylogenies. % (hemagglutinin and neuraminidase in green and magenta respectively).
} % END: footnotesize
}
\label{fig:discrete}
\end{figure}

@@

\begin{figure}[h!]
\centering
\includegraphics[scale=0.2]{moore}
\caption{
{ \footnotesize 
{\bf Performance gains of a numerical code benchmarked on a single CPU core over the course of 15 years (PDE solver FeatFlow, \citep{Turek1999}).}
Last two years have seen only a marginal speed-up, and correspond to the shift to multicore architectures.
} % END: footnotesize
}
\label{fig:moore}
\end{figure}
